A Hexagon Tile Map Algorithm for Displaying Spatial Data

### 2020-12-09 kobakian-cook.Rmd
Use RJournal commands for text marking

This paper contains a discussion of existing mapping practices. This is followed by details of the algorithm. The implementation section discusses the algorithm in the `sugarbag` package. Finally, the use and benefits of animation are presented.

@@

There are several key steps in the creation of the hexagon tile map as described in the flow chart in Figure \ref{fig:sugarbag_flow}. First, derive the set of centroids from the polygons provided, then create the grid of hexagon locations. These two processes are defined in the blue left column of the flow chart in Figure \ref{fig:sugarbag_flow}. Each centroid can then be allocated to an available hexagon location. The steps for the allocation process are detailed in the right column of Figure \ref{fig:sugarbag_flow}. There are several filter steps to speed the process of selecting an appropriate hexagon to represent each geographic unit. To make tessellated plots with the hexagon allocations, the point locations are converted into hexagon shapes.

@@

<!-- The`create_hexmap` function will -->

@@

A user can tweak the parameters of the hexagon map using the additional arguments to the `create_hexmap` function, but these options may affect the speed of the algorithm. 
The hexagon size may need adjustments depending on the density of the population; users can provide an appropriate hexagon size and re-run the algorithm. The buffer distance may need to be increased if the coastal cities need to extend beyond the geographic land mass.

@@

The following steps create the hexagon tile map for all the Statistical Areas at Level 2 in Tasmania. These steps can be executed by the main function, `create_hexmap`, or can be run separately for more flexibility.

If a user would like to perform each of the steps of the algorithm themselves, the necessary inputs will change for each function. For example, the `allocate` function can be used directly when a user wishes to use a set of centroids, rather than polygons.

@@

The buffer distance, hexagon size, hexagon amount to filter and width of angle are parameters that will be determined within `create_hexmap()`, if they are not provided. They are created as they are needed throughout the following example.

@@

### Step 1: Derive the set of centroid points

The set of polygons should be provided as an `sf` object, this is a data frame containing a `geometry` column. The `read_shape()` function can assist in creating this object for use in `R`.

The centroids can be derived from an `sf` object using the `create_centroids()` function:

@@

### Step 2: Create the hexagon grid points

A grid is created to allow tessellation of the hexagons that represent the geographic units. For a hexagon tile map, the grid of possible hexagon locations is made using the `create_grid()` function. It uses the centroids, the hexagon size and the buffer distance. Figure \ref{fig:filterprocess}a) shows the initial locations of grid points created for Tasmania.

#### Creating a tessellated grid

A set of longitude columns, and latitude rows are created to define the locations of the hexagons. The distance between each row and column is the size specified by `hex_size`. Equally spaced columns are created from the minimum longitude minus the buffer distance, up to the maximum longitude plus the buffer distance. Similarly, the rows are created from the latitude values and the buffer distance. A unique hexagon location is created from all intersections of the longitude columns and latitude rows. Figure \ref{fig:filterprocess}a shows the hexagon grid after every second latitude row on the grid is shifted right, by half of the hexagon size.

@@

#### Filtering the grid

@@

The hexagon grid points are kept only if they fall within the rolling average of the minimum and maximum centroid values after accounting for the buffer distance, for each row and column of the grid. Figure \ref{fig:filterprocess}b displays remaining hexagon grid points after applying the buffer filter. The sparsely populated South-West region of National Park has fewer points available in the water compared to the South-East region near the city of Hobart.

@@

The distance is calculated between each centroid in the set, and each of the focal points provided. The order for allocation is determined by the distance between the polygon centroid and it's closest focal point. For this example, this distance is only calculated for the capital city of Hobart, represented in Figure \ref{fig:filterprocess}c and d as a white cross.

### Step 3: Allocate each centroid to a hexagon grid point

@@

#### Filter the grid for unassigned hexagon points

The set of possible hexagon grid points reduces by one after each centroid is allocated, then only the grid points that have not yet been selected are considered. Keeping only the available hexagon points prevents multiple geographic units from being allocated to the same hexagon. This is demonstrated in Figure \ref{fig:filterprocess}c and d by the black hexagons that represent the seven closest polygons to the capital city of Hobart. As the allocation process begins for the eighth closest centroid there are seven unavailable hexagon locations.

#### Filter the grid points for those closest to the centroid

A selection of possible hexagon locations around the centroid allows only the closest points that are not yet assigned to be considered. The number of possible hexagon locations to consider for a centroid is determined by the `hex_filter`. This is the maximum amount of hexagons between the centroid and the furthest considered hexagon. It is used to subset possible grid points to only those surrounding the polygon centroid within an appropriate range. A smaller distance will increase speed, but can decrease accuracy when width of the angle increases.

@@

f1 <- ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = interaction(sa2_name_2011, polygon)), data = fort_tas, fill = "#D3D3D3", colour = "white", size = 0.1) +
  theme_void() + coord_equal(ylim = c(-44, -40.1), xlim = c(144.5, 149)) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#55b56c", data = grid, size = 0.5) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#09661f", data = f_grid, size = 1) +
  # allocated areas
    geom_polygon(aes(x = long, y = lat, group = interaction(sa2_name_2011)), data = fort_allocated, fill = "#353b38", colour = "black", size = 0.01, alpha = 0.7) +
  #Hobart
  geom_point(aes(x = focal_longitude, y = focal_latitude), data = f_centroid, colour = "#e9f2ed", size = 4, shape = 3) + 
  geom_point(aes(x=longitude, y = latitude), data = f_centroid, fill = "#beaed4", shape = 25)

@@

The `width` parameter is used to take a slice of the remaining points. The width is the amount of degrees used on either side of the angle from the focal point to centroid location. This uses the angle from the closest capital city, to the current centroid as seen in Figure \ref{fig:filterprocess}c. This allows the spatial relationship to be preserved, even when it is allocated to a hexagon that is further from the focal point then the original centroid location.

@@

f2 <- ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = interaction(sa2_name_2011, polygon)), data = fort_tas %>% filter(lat < -40.3), fill = "#D3D3D3", colour = "white", size = 0.1) +
  geom_polygon(aes(x = long, y = lat, group = interaction(sa2_name_2011)), data = fort_allocated, fill = "#353b38", colour = "black", size = 0.01, alpha = 0.7) +
  #Hobart
  geom_point(aes(x = focal_longitude, y = focal_latitude), data = f_centroid, colour = "#e9f2ed", size = 4, shape = 3) + 
  geom_point(aes(x=longitude, y = latitude), data = f_centroid, fill = "#beaed4", shape = 25, size = 2) + 
  theme_void() + coord_equal(xlim = c(146.9, 147.8), ylim = c(-43.35, -42.5)) +
  # possible choices to select
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#09661f", data = f_grid, size = 2.5) +
  # selected hexagon
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#09661f", fill = "#beaed4", data = hex, size = 3.5, shape = 21)

@@

cowplot::plot_grid(g1, g2,
                   f1, f2, labels = c("a", "b", "c", "d"))
```

If no available hexagon grid point is found within the original filter distance and angle, the distance is expanded and only when a maximum distance is reached will the angle expand to accommodate more possible grid points.
By default the angle filter for the hexagon grid points create plus and minus 30 degree bounds of the angle from the focal point to the geographic centroid. This will increase if no points can be found within the `hex_filter` distance. The default angle of 30 was chosen to allow the algorithm to find hexagons that best maintained the spatial relationship between the focal point and geographic centroid.

@@

Creating an animation connecting these two displays can allow users to grasp the meaning of the connected hexagons. It will also highlight the density of the communities using the rapid expansion of the inner-city areas, these hexagons will move the furthest and will move rapidly in the animation from their geographic locations. The rapid decrease in size of the large rural areas can show the large size of their collective landmass. The `gganimate` [@gganimate] package can be used to make an animation. It connects the polygons for each area in the two displays using the `sf_id` variable, the names of the statistical areas, and creates intermediate displays as they geographic areas shrink or expand to the chosen hexagon size.

@@

<!-- implementation -->
 The `sugarbag` package for `R` automates the creation of tessellated hexagon tile maps by providing an algorithm to design these displays. The Australian application preserves spatial relationships, and emphasises capital cities. However, similar to the choropleth map display the tessellation does not allow the size of the hexagons to represent another variable. The algorithm uses focal points, which are the centers of capital cities for Australia, to determine the order and direction of allocation. After careful consideration of the choropleth map, the small geographic inner city areas with high thyroid cancer prevalence may have been missed by viewers, but the hexagon tile map display reveals them. The communities in northern Tasmania and the Northern Territory draw less attention because they are smaller and more isolated in the hexagon tile map, but their colour is still comparable with the hexagons in other regions.

>>>>

This paper contains a discussion of Existing Mapping Practices. This is followed by details of the Algorithm. The implementation section discusses the algorithm in the \pkg{sugarbag} package. Finally, the use and benefits of animation are presented.

@@

There are several key steps in the creation of the hexagon tile map as described in the flow chart in Figure \ref{fig:sugarbag_flow}. First, derive the set of centroids from the polygons provided, then create the grid of hexagon locations. These two processes are defined in the blue left column of the flow chart in figure \ref{fig:sugarbag_flow}. Each centroid can then be allocated to an available hexagon location. The steps for the allocation process are detailed in the right column of Figure \ref{fig:sugarbag_flow}. There are several filter steps to speed the process of selecting an appropriate hexagon to represent each geographic unit. To make tessellated plots with the hexagon allocations, the point locations are converted into hexagon shapes.

@@

<!-- The\code{create\_hexmap} function will -->

@@

A user can tweak the parameters of the hexagon map using the additional arguments to the \code{create\_hexmap} function, but these options may affect the speed of the algorithm. 
The hexagon size may need adjustments depending on the density of the population; users can provide an appropriate hexagon size and re-run the algorithm. The buffer distance may need to be increased if the coastal cities need to extend beyond the geographic land mass.

@@

The following steps create the hexagon tile map for all the Statistical Areas at Level 2 in Tasmania. These steps can be executed by the main function, \code{create\_hexmap}, or can be run separately for more flexibility.

If a user would like to perform each of the steps of the algorithm themselves, the necessary inputs will change for each function. For example, the \code{allocate} function can be used directly when a user wishes to use a set of centroids, rather than polygons.

@@

The buffer distance, hexagon size, hexagon amount to filter and width of angle are parameters that will be determined within \code{create\_hexmap}, if they are not provided. They are created as they are needed throughout the following example.

@@

### Derive the set of centroid points

The set of polygons should be provided as an \pkg{sf} object, this is a data frame containing a \code{geometry} column. The \code{read\_shape} function can assist in creating this object for use in R.

The centroids can be derived from an \pkg{sf} object using the \code{create\_centroids} function:

@@

### Step 1: Create the hexagon grid points

A grid is created to allow tessellation of the hexagons that represent the geographic units. For a hexagon tile map, the grid of possible hexagon locations is made using the \code{create\_grid} function. It uses the centroids, the hexagon size and the buffer distance. Figure \ref{fig:filterprocess} 1a) shows the initial locations of grid points created for Tasmania.

#### (a) Creating a tessellated grid

A set of longitude columns, and latitude rows are created to define the locations of the hexagons. The distance between each row and column is the size specified by \code{hex\_size}. Equally spaced columns are created from the minimum longitude minus the buffer distance, up to the maximum longitude plus the buffer distance. Similarly, the rows are created from the latitude values and the buffer distance. A unique hexagon location is created from all intersections of the longitude columns and latitude rows. Figure \ref{fig:filterprocess} 1a) shows the hexagon grid after every second latitude row on the grid is shifted right, by half of the hexagon size.

@@

#### (b) Filtering the grid

Not all of the grid points will be used, especially if islands result in a large grid space. To filter the grid for appropriate hexagon locations for allocation, the \code{create\_grid} function will call the \code{create\_buffer} function. It finds the grid points needed to best capture the set of centroids on a hexagon tile map. Reducing the buffer can help to decrease the amount of time taken, as there will be less points over the water to consider for each centroid allocation.

@@

The hexagon grid points are kept only if they fall within the rolling average of the minimum and maximum centroid values after accounting for the buffer distance, for each row and column of the grid. Figure \ref{fig:filterprocess} 1b) displays remaining hexagon grid points after applying the buffer filter. The sparsely populated South-West region of National Park has fewer points available in the water compared to the South-East region near the city of Hobart.

@@

The distance is calculated between each centroid in the set, and each of the focal points provided. The order for allocation is determined by the distance between the polygon centroid and it's closest focal point. For this example, this distance is only calculated for the capital city of Hobart, represented in figure \ref{fig:filterprocess} 2a) and 2b) as a white cross.

### Step 2: Allocate each centroid to a hexagon grid point

@@

#### (a) Filter the grid for unassigned hexagon points

The set of possible hexagon grid points reduces by one after each centroid is allocated, then only the grid points that have not yet been selected are considered. Keeping only the available hexagon points prevents multiple geographic units from being allocated to the same hexagon. This is demonstrated in figure \ref{fig:filterprocess} 2a) and 2b) by the black hexagons that represent the seven closest polygons to the capital city of Hobart. As the allocation process begins for the eighth closest centroid there are seven unavailable hexagon locations.

#### (b) Filter the grid points for those closest to the centroid

A selection of possible hexagon locations around the centroid allows only the closest points that are not yet assigned to be considered. The number of possible hexagon locations to consider for a centroid is determined by the hexagon filter. This is the maximum amount of hexagons between the centroid and the furthest considered hexagon. It is used to subset possible grid points to only those surrounding the polygon centroid within an appropriate range. A smaller distance will increase speed, but can decrease accuracy when width of the angle increases.

@@

f1 <- ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = interaction(sa2_name_2011, polygon)), data = fort_tas, fill = "#D3D3D3", colour = "white", size = 0.1) +
  theme_void() + coord_equal(ylim = c(-45.1, -40.5), xlim = c(144.5, 149)) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#55b56c", data = grid, size = 0.5) +
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#09661f", data = f_grid, size = 1) +
  # allocated areas
    geom_polygon(aes(x = long, y = lat, group = interaction(sa2_name_2011)), data = fort_allocated, fill = "#353b38", colour = "black", size = 0.01, alpha = 0.7) +
  #Hobart
  geom_point(aes(x = focal_longitude, y = focal_latitude), data = f_centroid, colour = "#e9f2ed", size = 4, shape = 3) + 
  geom_point(aes(x=longitude, y = latitude), data = f_centroid, fill = "#beaed4", shape = 25)

@@

The \code{width} parameter is used to take a slice of the remaining points. The width is the amount of degrees used on either side of the angle from the focal point to centroid location. This uses the angle from the closest capital city, to the current centroid as seen in figure \ref{fig:filterprocess} 2a). This allows the spatial relationship to be preserved, even when it is allocated to a hexagon that is further from the focal point then the original centroid location.

@@

f2 <- ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = interaction(sa2_name_2011, polygon)), data = fort_tas %>% filter(lat < -40.3), fill = "#D3D3D3", colour = "white", size = 0.1) +
  geom_polygon(aes(x = long, y = lat, group = interaction(sa2_name_2011)), data = fort_allocated, fill = "#353b38", colour = "black", size = 0.01, alpha = 0.7) +
  #Hobart
  geom_point(aes(x = focal_longitude, y = focal_latitude), data = f_centroid, colour = "#e9f2ed", size = 4, shape = 3) + 
  geom_point(aes(x=longitude, y = latitude), data = f_centroid, fill = "#beaed4", shape = 25, size = 2) + 
  theme_void() + coord_equal(xlim = c(146.9, 148), ylim = c(-43.6, -42.5)) +
  # possible choices to select
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#09661f", data = f_grid, size = 2.5) +
  # selected hexagon
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#09661f", fill = "#beaed4", data = hex, size = 3.5, shape = 21)

@@

cowplot::plot_grid(g1, g2,
                   f1, f2,
                   labels = c("1a", "1b", "2a", "2b"))
```

If no available hexagon grid point is found within the original filter distance and angle, the distance is expanded and only when a maximum distance is reached will the angle expand to accommodate more possible grid points.
By default the angle filter for the hexagon grid points create plus and minus 30 degree bounds of the angle from the focal point to the geographic centroid. This will increase if no points can be found within the \code{hex\_filter} distance. The default angle of 30 was chosen to allow the algorithm to find hexagons that best maintained the spatial relationship between the focal point and geographic centroid.

@@

Creating an animation connecting these two displays can allow users to grasp the meaning of the connected hexagons. It will also highlight the density of the communities using the rapid expansion of the inner-city areas, these hexagons will move the furthest and will move rapidly in the animation from their geographic locations. The rapid decrease in size of the large rural areas can show the large size of their collective landmass. The \pkg{gganimate} [@gganimate] package can be used to make an animation. It connects the polygons for each area in the two displays using the \code{sf\_id} variable, the names of the statistical areas, and creates intermediate displays as they geographic areas shrink or expand to the chosen hexagon size.

@@

<!-- implementation -->

The hexagon tile map display acknowledges that the amount of residents can differ but each administrative area is equally important. This solution uses equally sized areas, and maintain neighbourhood boundary connections. The \pkg{sugarbag} package for R automates the creation of tessellated hexagon tile maps by providing an algorithm to design these displays. The Australian application preserves spatial relationships, and emphasises capital cities. 
However, similar to the choropleth map display the tessellation does not allow the size of the hexagons to represent another variable. The algorithm is heavily dependent on the focal points used, as this determines the order of allocation. With careful consideration of the choropleth map, the small geographic inner city areas may not have been noticed by viewers, but the hexagon tile map display emphasises them. The communities in northern Tasmania and the Northern territory do not draw attention because of their size as in the choropleth, but their colour is still noticeably below average when contrasted with the hexagons in New South Wales.

### 2020-12-09 kobakian-cook.tex
Use RJournal commands for text marking

This paper contains a discussion of existing mapping practices. This is
followed by details of the algorithm. The implementation section
discusses the algorithm in the \texttt{sugarbag} package. Finally, the
use and benefits of animation are presented.

@@

A user can tweak the parameters of the hexagon map using the additional
arguments to the \texttt{create\_hexmap} function, but these options may
affect the speed of the algorithm. The hexagon size may need adjustments
depending on the density of the population; users can provide an
appropriate hexagon size and re-run the algorithm. The buffer distance
may need to be increased if the coastal cities need to extend beyond the
geographic land mass.

@@

The following steps create the hexagon tile map for all the Statistical
Areas at Level 2 in Tasmania. These steps can be executed by the main
function, \texttt{create\_hexmap}, or can be run separately for more
flexibility.

If a user would like to perform each of the steps of the algorithm
themselves, the necessary inputs will change for each function. For
example, the \texttt{allocate} function can be used directly when a user
wishes to use a set of centroids, rather than polygons.

@@

The buffer distance, hexagon size, hexagon amount to filter and width of
angle are parameters that will be determined within
\texttt{create\_hexmap()}, if they are not provided. They are created as
they are needed throughout the following example.

@@

\hypertarget{step-1-derive-the-set-of-centroid-points}{%
\subsubsection{Step 1: Derive the set of centroid
points}\label{step-1-derive-the-set-of-centroid-points}}

The set of polygons should be provided as an \texttt{sf} object, this is
a data frame containing a \texttt{geometry} column. The
\texttt{read\_shape()} function can assist in creating this object for
use in \texttt{R}.

The centroids can be derived from an \texttt{sf} object using the
\texttt{create\_centroids()} function:

\hypertarget{step-2-create-the-hexagon-grid-points}{%
\subsubsection{Step 2: Create the hexagon grid
points}\label{step-2-create-the-hexagon-grid-points}}

A grid is created to allow tessellation of the hexagons that represent
the geographic units. For a hexagon tile map, the grid of possible
hexagon locations is made using the \texttt{create\_grid()} function. It
uses the centroids, the hexagon size and the buffer distance. Figure
\ref{fig:filterprocess}a) shows the initial locations of grid points
created for Tasmania.

\hypertarget{creating-a-tessellated-grid}{%
\paragraph{Creating a tessellated
grid}\label{creating-a-tessellated-grid}}

A set of longitude columns, and latitude rows are created to define the
locations of the hexagons. The distance between each row and column is
the size specified by \texttt{hex\_size}. Equally spaced columns are
created from the minimum longitude minus the buffer distance, up to the
maximum longitude plus the buffer distance. Similarly, the rows are
created from the latitude values and the buffer distance. A unique
hexagon location is created from all intersections of the longitude
columns and latitude rows. Figure \ref{fig:filterprocess}a shows the
hexagon grid after every second latitude row on the grid is shifted
right, by half of the hexagon size.

\hypertarget{filtering-the-grid}{%
\paragraph{Filtering the grid}\label{filtering-the-grid}}

@@

The distance is calculated between each centroid in the set, and each of
the focal points provided. The order for allocation is determined by the
distance between the polygon centroid and it's closest focal point. For
this example, this distance is only calculated for the capital city of
Hobart, represented in Figure \ref{fig:filterprocess}c and d as a white
cross.

\hypertarget{step-3-allocate-each-centroid-to-a-hexagon-grid-point}{%
\subsubsection{Step 3: Allocate each centroid to a hexagon grid
point}\label{step-3-allocate-each-centroid-to-a-hexagon-grid-point}}

@@

\hypertarget{filter-the-grid-for-unassigned-hexagon-points}{%
\paragraph{Filter the grid for unassigned hexagon
points}\label{filter-the-grid-for-unassigned-hexagon-points}}

The set of possible hexagon grid points reduces by one after each
centroid is allocated, then only the grid points that have not yet been
selected are considered. Keeping only the available hexagon points
prevents multiple geographic units from being allocated to the same
hexagon. This is demonstrated in Figure \ref{fig:filterprocess}c and d
by the black hexagons that represent the seven closest polygons to the
capital city of Hobart. As the allocation process begins for the eighth
closest centroid there are seven unavailable hexagon locations.

\hypertarget{filter-the-grid-points-for-those-closest-to-the-centroid}{%
\paragraph{Filter the grid points for those closest to the
centroid}\label{filter-the-grid-points-for-those-closest-to-the-centroid}}

A selection of possible hexagon locations around the centroid allows
only the closest points that are not yet assigned to be considered. The
number of possible hexagon locations to consider for a centroid is
determined by the \texttt{hex\_filter}. This is the maximum amount of
hexagons between the centroid and the furthest considered hexagon. It is
used to subset possible grid points to only those surrounding the
polygon centroid within an appropriate range. A smaller distance will
increase speed, but can decrease accuracy when width of the angle
increases.

@@

The \texttt{width} parameter is used to take a slice of the remaining
points. The width is the amount of degrees used on either side of the
angle from the focal point to centroid location. This uses the angle
from the closest capital city, to the current centroid as seen in Figure
\ref{fig:filterprocess}c.~This allows the spatial relationship to be
preserved, even when it is allocated to a hexagon that is further from
the focal point then the original centroid location.

@@

If no available hexagon grid point is found within the original filter
distance and angle, the distance is expanded and only when a maximum
distance is reached will the angle expand to accommodate more possible
grid points. By default the angle filter for the hexagon grid points
create plus and minus 30 degree bounds of the angle from the focal point
to the geographic centroid. This will increase if no points can be found
within the \texttt{hex\_filter} distance. The default angle of 30 was
chosen to allow the algorithm to find hexagons that best maintained the
spatial relationship between the focal point and geographic centroid.

@@

Creating an animation connecting these two displays can allow users to
grasp the meaning of the connected hexagons. It will also highlight the
density of the communities using the rapid expansion of the inner-city
areas, these hexagons will move the furthest and will move rapidly in
the animation from their geographic locations. The rapid decrease in
size of the large rural areas can show the large size of their
collective landmass. The \texttt{gganimate} \citep{gganimate} package
can be used to make an animation. It connects the polygons for each area
in the two displays using the \texttt{sf\_id} variable, the names of the
statistical areas, and creates intermediate displays as they geographic
areas shrink or expand to the chosen hexagon size.

@@

The \texttt{sugarbag} package for \texttt{R} automates the creation of
tessellated hexagon tile maps by providing an algorithm to design these
displays. The Australian application preserves spatial relationships,
and emphasises capital cities. However, similar to the choropleth map
display the tessellation does not allow the size of the hexagons to
represent another variable. The algorithm uses focal points, which are
the centers of capital cities for Australia, to determine the order and
direction of allocation. After careful consideration of the choropleth
map, the small geographic inner city areas with high thyroid cancer
prevalence may have been missed by viewers, but the hexagon tile map
display reveals them. The communities in northern Tasmania and the
Northern Territory draw less attention because they are smaller and more
isolated in the hexagon tile map, but their colour is still comparable
with the hexagons in other regions.

>>>>

This paper contains a discussion of Existing Mapping Practices. This is
followed by details of the Algorithm. The implementation section
discusses the algorithm in the \pkg{sugarbag} package. Finally, the use
and benefits of animation are presented.

@@

A user can tweak the parameters of the hexagon map using the additional
arguments to the \code{create\_hexmap} function, but these options may
affect the speed of the algorithm. The hexagon size may need adjustments
depending on the density of the population; users can provide an
appropriate hexagon size and re-run the algorithm. The buffer distance
may need to be increased if the coastal cities need to extend beyond the
geographic land mass.

@@

The following steps create the hexagon tile map for all the Statistical
Areas at Level 2 in Tasmania. These steps can be executed by the main
function, \code{create\_hexmap}, or can be run separately for more
flexibility.

If a user would like to perform each of the steps of the algorithm
themselves, the necessary inputs will change for each function. For
example, the \code{allocate} function can be used directly when a user
wishes to use a set of centroids, rather than polygons.

@@

The buffer distance, hexagon size, hexagon amount to filter and width of
angle are parameters that will be determined within
\code{create\_hexmap}, if they are not provided. They are created as
they are needed throughout the following example.

@@

\hypertarget{derive-the-set-of-centroid-points}{%
\subsubsection{Derive the set of centroid
points}\label{derive-the-set-of-centroid-points}}

The set of polygons should be provided as an \pkg{sf} object, this is a
data frame containing a \code{geometry} column. The \code{read\_shape}
function can assist in creating this object for use in R.

The centroids can be derived from an \pkg{sf} object using the
\code{create\_centroids} function:

\hypertarget{step-1-create-the-hexagon-grid-points}{%
\subsubsection{Step 1: Create the hexagon grid
points}\label{step-1-create-the-hexagon-grid-points}}

A grid is created to allow tessellation of the hexagons that represent
the geographic units. For a hexagon tile map, the grid of possible
hexagon locations is made using the \code{create\_grid} function. It
uses the centroids, the hexagon size and the buffer distance. Figure
\ref{fig:filterprocess} 1a) shows the initial locations of grid points
created for Tasmania.

\hypertarget{a-creating-a-tessellated-grid}{%
\paragraph{(a) Creating a tessellated
grid}\label{a-creating-a-tessellated-grid}}

A set of longitude columns, and latitude rows are created to define the
locations of the hexagons. The distance between each row and column is
the size specified by \code{hex\_size}. Equally spaced columns are
created from the minimum longitude minus the buffer distance, up to the
maximum longitude plus the buffer distance. Similarly, the rows are
created from the latitude values and the buffer distance. A unique
hexagon location is created from all intersections of the longitude
columns and latitude rows. Figure \ref{fig:filterprocess} 1a) shows the
hexagon grid after every second latitude row on the grid is shifted
right, by half of the hexagon size.

\hypertarget{b-filtering-the-grid}{%
\paragraph{(b) Filtering the grid}\label{b-filtering-the-grid}}

Not all of the grid points will be used, especially if islands result in
a large grid space. To filter the grid for appropriate hexagon locations
for allocation, the \code{create\_grid} function will call the
\code{create\_buffer} function. It finds the grid points needed to best
capture the set of centroids on a hexagon tile map. Reducing the buffer
can help to decrease the amount of time taken, as there will be less
points over the water to consider for each centroid allocation.

@@

The distance is calculated between each centroid in the set, and each of
the focal points provided. The order for allocation is determined by the
distance between the polygon centroid and it's closest focal point. For
this example, this distance is only calculated for the capital city of
Hobart, represented in figure \ref{fig:filterprocess} 2a) and 2b) as a
white cross.

\hypertarget{step-2-allocate-each-centroid-to-a-hexagon-grid-point}{%
\subsubsection{Step 2: Allocate each centroid to a hexagon grid
point}\label{step-2-allocate-each-centroid-to-a-hexagon-grid-point}}

@@

\hypertarget{a-filter-the-grid-for-unassigned-hexagon-points}{%
\paragraph{(a) Filter the grid for unassigned hexagon
points}\label{a-filter-the-grid-for-unassigned-hexagon-points}}

The set of possible hexagon grid points reduces by one after each
centroid is allocated, then only the grid points that have not yet been
selected are considered. Keeping only the available hexagon points
prevents multiple geographic units from being allocated to the same
hexagon. This is demonstrated in figure \ref{fig:filterprocess} 2a) and
2b) by the black hexagons that represent the seven closest polygons to
the capital city of Hobart. As the allocation process begins for the
eighth closest centroid there are seven unavailable hexagon locations.

\hypertarget{b-filter-the-grid-points-for-those-closest-to-the-centroid}{%
\paragraph{(b) Filter the grid points for those closest to the
centroid}\label{b-filter-the-grid-points-for-those-closest-to-the-centroid}}

A selection of possible hexagon locations around the centroid allows
only the closest points that are not yet assigned to be considered. The
number of possible hexagon locations to consider for a centroid is
determined by the hexagon filter. This is the maximum amount of hexagons
between the centroid and the furthest considered hexagon. It is used to
subset possible grid points to only those surrounding the polygon
centroid within an appropriate range. A smaller distance will increase
speed, but can decrease accuracy when width of the angle increases.

@@

The \code{width} parameter is used to take a slice of the remaining
points. The width is the amount of degrees used on either side of the
angle from the focal point to centroid location. This uses the angle
from the closest capital city, to the current centroid as seen in figure
\ref{fig:filterprocess} 2a). This allows the spatial relationship to be
preserved, even when it is allocated to a hexagon that is further from
the focal point then the original centroid location.

@@

If no available hexagon grid point is found within the original filter
distance and angle, the distance is expanded and only when a maximum
distance is reached will the angle expand to accommodate more possible
grid points. By default the angle filter for the hexagon grid points
create plus and minus 30 degree bounds of the angle from the focal point
to the geographic centroid. This will increase if no points can be found
within the \code{hex\_filter} distance. The default angle of 30 was
chosen to allow the algorithm to find hexagons that best maintained the
spatial relationship between the focal point and geographic centroid.

@@

Creating an animation connecting these two displays can allow users to
grasp the meaning of the connected hexagons. It will also highlight the
density of the communities using the rapid expansion of the inner-city
areas, these hexagons will move the furthest and will move rapidly in
the animation from their geographic locations. The rapid decrease in
size of the large rural areas can show the large size of their
collective landmass. The \pkg{gganimate} \citep{gganimate} package can
be used to make an animation. It connects the polygons for each area in
the two displays using the \code{sf\_id} variable, the names of the
statistical areas, and creates intermediate displays as they geographic
areas shrink or expand to the chosen hexagon size.

@@

The hexagon tile map display acknowledges that the amount of residents
can differ but each administrative area is equally important. This
solution uses equally sized areas, and maintain neighbourhood boundary
connections. The \pkg{sugarbag} package for R automates the creation of
tessellated hexagon tile maps by providing an algorithm to design these
displays. The Australian application preserves spatial relationships,
and emphasises capital cities. However, similar to the choropleth map
display the tessellation does not allow the size of the hexagons to
represent another variable. The algorithm is heavily dependent on the
focal points used, as this determines the order of allocation. With
careful consideration of the choropleth map, the small geographic inner
city areas may not have been noticed by viewers, but the hexagon tile
map display emphasises them. The communities in northern Tasmania and
the Northern territory do not draw attention because of their size as in
the choropleth, but their colour is still noticeably below average when
contrasted with the hexagons in New South Wales.
